### 2016-02-10 abstract.tex
minor fixes

When are two program executions similar?  In most contexts, the answer is that they are similar when they satisfy or violate correctness properties for the same reasons.  This paper proposes a novel metric for measuring the distance between two failing test cases, and automated debugging methods based on that metric.  The metric is based on a causal concept of distance:  executions are similar to the degree that changes in the program itself, called mutants, cause similar changes in the correctness of the executions.  Specifically, if two failing test cases become successful due to the same mutant, they are more likely to be due to the same fault.  We evaluate the mutant-based distance metric by applying it to problems of fault identification and automated debugging assistance for two widely-used real-world compilers, and demonstrate improvements over state-of-the-art methods.

>>>>

When are two program executions similar?  In most contexts, the answer is that they are similar when they satisfy or violate correctness properties for the same reasons.  This paper proposes a novel metric for measuring the distance between two failing test cases, and automated debugging methods based on that metric.  The metric is based on a causal concept of distance:  executions are similar to the degree that changes in the program itself, introduced using mutation operators, cause similar changes in the correctness of the executions.  Specifically, if two failing test cases become successful due to the same mutant, they are more likely to be due to the same fault.  We evaluate the mutant-based distance metric by applying it to problems of fault identification and automated debugging assistance for two widely-used real-world compilers, and demonstrate improvements over state-of-the-art methods.

### 2016-02-10 conclusions.tex
minor fixes

%We also plan to explore other applications of mutant-based metrics.
Many software engineering techniques rely on measuring distance
between program executions.  Not all of these concern failing
executions.  For example, Zhang et al. \cite{issta14} show that FPF based on a simple
Hamming distance over branches covered can significantly improve the
effectiveness of seeded symbolic execution
\cite{Zesti,PersonSeed,BugRedux}.  Mutation response is likely too
expensive to use for this purpose without modification, but sampling
mutants to refine distance when distances plateau
may be practical.  Mutants may also be useful in determining
execution peers \cite{Sumner2011}, for example to help operators of
spacecraft find similar behaviors to telemetry downloads in testbed
history \cite{KlausRajeev,scriptstospecs}.

>>>>

%We also plan to explore other applications of mutant-based metrics.
Many software engineering techniques rely on measuring distance
between program executions.  Not all of these concern failing
executions.  For example, Zhang et al. \cite{issta14} show that FPF based on a simple
Hamming distance over branches covered can significantly improve the
effectiveness of seeded symbolic execution
\cite{Zesti,PersonSeed,BugRedux}.  Mutation response is likely too
expensive to use for this purpose without modification, but sampling
mutants to refine distance when coarser metrics plateau
may be practical.  Mutants may also be useful in determining
execution peers \cite{Sumner2011}, for example to help operators of
spacecraft find similar behaviors to telemetry downloads in testbed
history \cite{KlausRajeev,scriptstospecs}.

### 2016-02-10 experiments.tex
minor fixes

In one sense, the discovery curve improvements here are practically useful, but not spectacular.  The SpiderMonkey curve has an APFD only slightly more than 2\% better than the best result from previous FPF efforts.  The GCC wrong code curve improves APFD by 6.4\%.  However, this is comparing the best curves from many curves computed in previous work to a single curve using a universal metric.  The comparison ignores the human effort required to find a good metric for any given program using just the techniques of our previous fuzzer taming work.

For fault localization and error explanation, the results show possible improvement on state-of-the-art approaches, in the absence of complete information from mutant analysis over passing tests (which can be very expensive to obtain).  A practical impact of the results reported is that we suggest users of our techniques examine only the first few (at most 10, and we propose as few as 5) mutants and coverage-changing statements, and ignore localization if none of these results are helpful.  This is analogous to our suggestion that a user abandon the FPF curve if 10 test cases in a row fail to reveal any new faults, as a heuristic.  We suspect the information in the mutants/coverage changes is probably helpful in some cases where they do not localize a faulty line, but this is simply based on our highly incomplete understanding of the faults and patches in question, and not a solidly established claim.  The expertise of compiler developers would be needed to confirm or reject this belief.  Our suggestion that mutations themselves provide interesting error explanation is also applicable to MUSE and MUSEUM.  There may be some potential for confusing explanations, however, if repairs include mutants that also cause some failures, as in the standard MUSE/MUSEUM approach.

One obvious question is:  why does {\bf Repair} perform best in this setting?  It is impossible to be sure, but one  possibility is power to distinguish repairs. If each failure was only repaired by one mutant, MUSEUM and {\bf Repair} would both rank that mutant maximally, and we speculate it would usually be faulty.  But such cases are rare:  for SpiderMonkey the mean/median number of mutants repairing each failure were 90.8 and 78, respectively, and for GCC the mean and median were 54 and 43.  If there are many repairs for a failure, how can the most important ones be distinguished?  MUSEUM relies on a large set of mutation operators, and assumes that faulty statements will have multiple repairs, compared to non-faulty statements, or that faulty statements will be less likely to cause passing tests to fail when mutated.  Restricting ourselves to cases where {\bf Repair} and MUSEUM both provided a localization, we know that some faulty statement was in a repairing mutant, but that passing tests could not distinguish this mutant  from other repairs.  Many of the faulty statement repairs (as expected for optimizing compilers) were statement deletions and conditional negations.  For these, it seems unlikely that Proteum would produce more mutants repairing the statement, though perhaps Proteum would create repairs for other statements that cannot be modified into a repair by any of Andrews' operators.  {\bf Repair} can still distinguish such mutants by relying on the onion-ring structure:  when a repair also repairs test cases that otherwise do not resemble the failure being localized (in terms of its repairing mutants), {\bf Repair} assumes that repair is general to many different faults.  The hypothesis behind {\bf Repair} is that \emph{repairs specific to the fault causing a failure are most likely to modify code involved in the fault itself}.  This seems to be true, when such a repair exists.  If any repair for a failure modified a faulty statement, {\bf Repair} always ranked it in the top 34 localizations, ranked it in the top 5 in 6 of 10 such cases, and 3 times ranked it 1st.  MUSEUM, in contrast, ranked one such repair 195th, never ranked a fault in its top 3 statements, and only ranked a fault in its top 5 statements twice.  Of course, MUSEUM might be suffering from a lack of mutation operators, but in some cases this appears unlikely due to the nature of the repairs.

>>>>

In one sense, the discovery curve improvements here are practically useful, but not spectacular.  The SpiderMonkey curve has an APFD only slightly more than 2\% better than the best result from previous FPF efforts.  The GCC wrong code curve improves APFD by 6.4\%.  However, the comparison is with the very best curve chosen after running more than 16 different metrics.  In practice, users do not know ground truth to rank competing curves, so this is not a practical approach to fuzzer taming.  The best methods for different subjects in previous work varied widely, even in such peculiar ways as less-fine-grained coverage providing better results in some cases, worse in other cases.   In practice, previous work established that FPF could produce good curves, but gave very little useful guidance for choosing a metric for actual use of the technique, and included methods such as comparing output signatures that are inapplicable to many settings.  This paper presents a single curve using a universal metric, and achieves a 2-6\% percentage point improvement over the best of more than 16 different metrics studied in the previous work.

For fault localization and error explanation, the results show possible improvement on state-of-the-art approaches, in the absence of complete information from mutant analysis over passing tests (which can be very expensive to obtain).  A practical impact of the results reported is that we suggest users of our techniques examine only the first few (at most 10, and we propose as few as 5) mutants and coverage-changing statements, and ignore localization if none of these results are helpful.  This is analogous to our suggestion that a user abandon the FPF curve if 5-10 test cases in a row fail to reveal any new faults, as a heuristic.  We suspect the information in the mutants/coverage changes is probably helpful in some cases where they do not localize a faulty line, but this is simply based on our highly incomplete understanding of the faults and patches in question, and not a solidly established claim.  The expertise of compiler developers would be needed to confirm or reject this belief.  Our suggestion that mutations themselves provide interesting error explanation is also applicable to MUSE and MUSEUM.  There may be some potential for confusing explanations, however, if repairs include mutants that also cause some failures, as in the standard MUSE/MUSEUM approach.

One obvious question is:  why does {\bf Repair} perform best in this setting?  It is impossible to be sure, but one  possibility is power to distinguish repairs. If each failure was only repaired by one mutant, MUSEUM and {\bf Repair} would both rank that mutant maximally, and we speculate it would usually be faulty.  But such cases are rare:  for SpiderMonkey the mean/median number of mutants repairing each failure were 90.8 and 78, respectively, and for GCC the mean and median were 54 and 43.  If there are many repairs for a failure, how can the most important ones be distinguished?  MUSEUM relies on a large set of mutation operators, and assumes that faulty statements will have multiple repairs, compared to non-faulty statements, or that faulty statements will be less likely to cause passing tests to fail when mutated.  Restricting ourselves to cases where {\bf Repair} and MUSEUM both provided a localization, we know that some faulty statement was in a repairing mutant, but that passing tests could not distinguish this mutant  from other repairs.  Many of the faulty statement repairs (as expected for optimizing compilers) were statement deletions and conditional negations.  For these, it seems unlikely that Proteum would produce more mutants repairing the statement, though perhaps Proteum would create repairs for other statements that cannot be modified into a repair by any of Andrews' operators.  {\bf Repair} can still distinguish such mutants by relying on the onion-ring structure:  when a repair also repairs test cases that otherwise do not resemble the failure being localized (in terms of its repairing mutants), {\bf Repair} assumes that repair is general to many different faults.  The hypothesis behind {\bf Repair} is that \emph{repairs specific to the fault causing a failure are most likely to modify code involved in the fault itself}.  This seems to be true, when such a repair exists.  If any repair for a failure modified a faulty statement, {\bf Repair} always ranked it in the top 34 localizations, ranked it in the top 5 in 6 of 10 such cases, and 3 times ranked it 1st.  MUSEUM, in contrast, ranked one such repair 195th, never ranked a fault in its top 3 statements, and only ranked a fault in its top 5 statements twice.  Of course, MUSEUM might be suffering from a lack of mutation operators, but in some cases this is unlikely due to the nature of the repairs.

### 2016-02-10 relatedwork.tex
minor fixes

The most closely related work is that of Chen et. al \cite{PLDI13}.  The present paper uses two of the same data sets, applies the FPF \cite{Gonzalez} algorithm, and has the same empirical focus of determining how to identify faults for complex compilers.  The primary difference is that the distance metrics used with FPF in the previous work were \emph{ad hoc}, chosen for each program by hand, produced poorer results for the critical portion of the curve for the two hard data sets, and were not extended to provide any fault localization.  This paper expands on Chen et al. by putting the FPF approach in the context of a much richer causal comparison of program executions that applies in theory to any executions, not just failing test cases.  Prior to the FPF-based work, Francis et al. \cite{Podgurski04}, Podgurski et al. \cite{Podgurski03} and others \cite{Liu06,Liblit05},used clustering to attempt to solve similar problems in identifying distinct faults, on simpler human-generated data sets.  The work of Chen et. al \cite{PLDI13} provides an in-depth comparison of clustering and ranking, and the advantages of ranking (as well as some possible limitations).  Jones et al. discuss general issues in debugging multiple faults efficiently \cite{Jones07}.

Fault localization \cite{FaultSurvey,NearNeighbor,Liblit03,Liblit05,Cleve05,Jones2002,Jones05,Jones07,GroceError,ChakiLev,Liu06,SPIN03,Santelices:ICSE:2009,Abreu:2006:PRDC,xuan2014test,DD}, even specifically for compilers \cite{Whalley94}, is a long-standing field of research.  Recently, fault localization researchers have used program mutants to improve statistical \cite{Jones2002} fault localization techniques \cite{MUSE,DebroyMutant,FasterMutant,multilingual,Metallaxis}.
%The core of this work has been the belief that faulty program locations are (1) more likely to change a failing test to successful when mutated and (2) less likely to change a successful test case to faulty when mutated.  This clearly connects to our general notion that how two test cases respond to a set of mutants is a good measure of their semantic similarity.
The work of Hong et al. \cite{multilingual} on MUSEUM and Moon et al. \cite{MUSE} on MUSE provides a good summary of other work along these lines, and is perhaps most similar to ours in assumptions, in particular ``Conjecture 1'' that a test case that used to fail on a program is more likely to pass on a mutation of a faulty statement.  While we suggest some localizations as a welcome and interesting side effect of our distance metric and FPF analysis, our primary goal is the discovery of all faults. Much fault localization work is largely (in many cases exclusively) based on single-fault scenarios; e.g., the MUSE \cite{MUSE} evaluation includes only 2 of 14 scenarios with multiple faults, and those scenarios include only a few faults.   The statistical approaches may owe some of their superiority in experiments over other methods to the use of mostly single-fault evaluations \cite{Jones05}.  By focusing on a single failing test case (for us, selected by fuzzer taming) as the locus of analysis, our work, like some early efforts \cite{NearNeighbor,Cleve05,GroceError} may be more suited for situations involving many faults.  MUSEUM \cite{multilingual} also revisits this approach, combining the MUSE formula with localization using only a single fault.

@@

The use of distance metrics in software engineering includes some of the localization and fault identification efforts discussed above.  Vangala et al. proposed using distance to cluster test cases to improve diversity and eliminate duplicates \cite{VangalaDist}.  Adaptive random testing uses distances (usually between inputs) to choose tests \cite{Chen,ARTChen,ISSTAART}, and Artzi et al. guided test generation for fault localization using path constraint metrics \cite{ArtziDirected}.  Sumner et al. evaluated approaches to generating execution peers for debugging and understanding \cite{Sumner2011} using tree edit distances (and evaluated other distances).  Xin et al. \cite{Xin2008} examined methods for indexing program executions, useful in all methods involving execution alignments.  To our knowledge, most metrics used are essentially spectrum-based \cite{RepsSpectra} (that is, using counts of coverage of code entities).    Our work differs in using a metric that is essentially causal, based on counterfactual \cite{LewisCause,LewisCount} programs.

>>>>

The most closely related work is that of Chen et. al \cite{PLDI13}.  The present paper uses two of the same data sets, applies the FPF \cite{Gonzalez} algorithm, and has the same empirical focus of determining how to identify faults for complex compilers.  The primary difference is that the distance metrics used with FPF in the previous work were \emph{ad hoc}, chosen for each program by hand, produced poorer results for the critical portion of the curve for the two hard data sets, and were not extended to provide any fault localization.  This paper expands on Chen et al. by putting the FPF approach in the context of a much richer causal comparison of program executions that applies in theory to any executions, not just failing test cases.  Prior to the FPF-based work, Francis et al. \cite{Podgurski04}, Podgurski et al. \cite{Podgurski03} and others \cite{Liu06,Liblit05} used clustering to attempt to solve similar problems in identifying distinct faults, on simpler human-generated data sets.  The work of Chen et. al \cite{PLDI13} provides an in-depth comparison of clustering and ranking, and the advantages of ranking (as well as some possible limitations).  Jones et al. discuss general issues in debugging multiple faults efficiently \cite{Jones07}.

Fault localization \cite{FaultSurvey,NearNeighbor,Liblit03,Liblit05,Cleve05,Jones2002,Jones05,Jones07,GroceError,ChakiLev,Liu06,SPIN03,Santelices:ICSE:2009,Abreu:2006:PRDC,xuan2014test,DD}, even specifically for compilers \cite{Whalley94}, is a long-standing field of research.  Recently, fault localization researchers have used program mutants to improve statistical \cite{Jones2002} fault localization techniques \cite{MUSE,DebroyMutant,FasterMutant,multilingual,Metallaxis}.
%The core of this work has been the belief that faulty program locations are (1) more likely to change a failing test to successful when mutated and (2) less likely to change a successful test case to faulty when mutated.  This clearly connects to our general notion that how two test cases respond to a set of mutants is a good measure of their semantic similarity.
The work of Hong et al. \cite{multilingual} on MUSEUM and Moon et al. \cite{MUSE} on MUSE provides a good summary of other work along these lines, and is perhaps most similar to ours in assumptions, in particular ``Conjecture 1'' that a test case that used to fail on a program is more likely to pass on a mutation of a faulty statement.  While we suggest some localizations as a welcome and interesting side effect of our distance metric and FPF analysis, our primary goal is the discovery of all faults. Much fault localization work is largely (in many cases exclusively) based on single-fault scenarios; e.g., the MUSE \cite{MUSE} evaluation includes only 2 of 14 scenarios with multiple faults, and those scenarios include only a few faults.   The statistical approaches may owe some of their superiority in experiments over other methods to the use of mostly single-fault evaluations \cite{Jones05}.  By focusing on a single failing test case (for us, selected by fuzzer taming) as the locus of analysis, our work, like some early efforts \cite{NearNeighbor,Cleve05,GroceError} may be more suited for situations involving many faults.  MUSEUM \cite{multilingual} also revisits this approach, combining the MUSE formula with localization using only a single failure.

@@

The use of distance metrics in software engineering includes some of the localization and fault identification efforts discussed above.  Vangala et al. proposed using distance to cluster test cases to improve diversity and eliminate duplicates \cite{VangalaDist}.  Adaptive random testing uses distances (usually between inputs) to choose tests \cite{Chen,ARTChen,ISSTAART}, and Artzi et al. guided test generation for fault localization using path constraint metrics \cite{ArtziDirected}.  Sumner et al. evaluated approaches to generating execution peers for debugging and understanding \cite{Sumner2011} using tree edit distances (and evaluated other distances).  Xin et al. \cite{Xin2008} examined methods for indexing program executions, useful in all methods involving execution alignments.  To our knowledge, most metrics used are essentially spectrum-based \cite{RepsSpectra} (using counts of coverage of code entities), while our metric is essentially causal, based on counterfactual \cite{LewisCause,LewisCount} programs.
